---
title: "Example TESAsamSim Run and DLM fit"
author: "Carrie Holt and Brendan Connors"
date: "January 26, 2021"
output:
  html_document: default
  pdf_document: default
---

This document illustrates how to simulate some spawner-recruitment data using the `TESAsamSim` R package and then fit a linear state-space model (aka Dynamic Linear model) with potentially time-varying parameters. The model is fit via Maximum likelihood followed by Kalman filtering and smoothing using a wrapper for the `dlm` R package (see [here](https://cran.r-project.org/web/packages/dlm/index.html)). For this illustrative example the simulation is based on Interior Fraser River coho salmon.


First install necessary packages and load Interior Fraser Coho data. 
```{r importData, echo=TRUE, warning=FALSE}
## Check if necessary packages are available and install if necessary
listOfPackages <- c("here", "parallel", "doParallel", "foreach", 
                    "tidyverse", "tictoc", "TESAsamSim")
newPackages <- listOfPackages[!(listOfPackages %in% 
                                  installed.packages()[ , "Package"])]
if (length(newPackages)) {
  install.packages(newPackages)
}
lapply(listOfPackages, require, character.only = TRUE)

# load wrapper function to fit dynamic linear models
source(here("dlm-wrapper.R"))

## Load relevant input data
# Simulation run parameters describing different scenarios
simPar <- read.csv(here("data", "IFCohoPars",
                         "cohoSimPar.csv"),
                       stringsAsFactors = F)
# CU-specific parameters
cuPar <- read.csv(here("data", "IFCohoPars", "cohoCUpars.csv"),
                  stringsAsFactors=F)
# Stock-recruit and catch data that are used to populate the simulation priming
# period
srDat <- read.csv(here("data", "IFCohoPars", "cohoRecDatTrim.csv"),
                  stringsAsFactors=F)

scenNames <- unique(simPar$scenario)
dirNames <- sapply(scenNames, function(x) paste(x, unique(simPar$species),
                                                sep = "_"))
```

Next, simulate some spawner-recruitment data, based on the known parameters specified in `cuPar` using the `genericRecoverySim()` function. For our illustrative purposes we only need one Monte Carlo trial, but simulator needs to run at least two.

```{r runSim, echo=TRUE, warning=FALSE}
genericRecoverySim(simPar=simPar[1,], cuPar=cuPar,  srDat=srDat,
            variableCU=FALSE, ricPars=NULL, #ricPars,  #cuCustomCorrMat = cuCustomCorrMat,
            dirName="example", nTrials=2, makeSubDirs=FALSE, random=FALSE)

```

In this example the outputs of the simulation are stored in outputs/simData/example. So let's load them and, since the simulator generates simulated data for 4 populations (CUs) and 2 Monte Carlo trials lets just grab the output from one of them.

```{r loadSimData, echo=TRUE, warning=FALSE}
simData <- read.csv(here("outputs", "simData", "example",
                         "ref_medFixedER_CUsrDat.csv"),
                       stringsAsFactors = F)

simDataCU1 <- simData %>% 
  filter(CU==1 & iteration ==1)

```
Now we can try fitting a Dynamic Linear Model to the simulated spawner-recruitment data. The model in this case is a linearized Ricker stock-recruit model with log(recruits/spawner) as the dependent variable, log(recruits/spawner) at low spawner abundance as the intercept (alpha), and a density dependence term as the slope (beta). To fit the model we will use a function we wrote for the workshop `fitDLM()` which calls functions from the `dlm`package. We can specify time-varying alpha and or beta with TRUE/FALSE statements in the function, which determine whether or not either parameter is modeled as a dynamic latent state. For now we will set the model to estimate time-varying alpha but not beta, even thought the simulated data are based on a stationary alpha and beta.

```{r fitDLM, echo=TRUE, warning=FALSE}
# rename a couple of columns to play nice with fitDLM. 
colnames(simDataCU1)[c(2,4,5,8,9)] <- c("byr","spwn", "rec", "alpha_true", "beta_true")

# fit DLM
dlm_model <- fitDLM(data = simDataCU1,
              alpha_vary = TRUE,
              beta_vary = FALSE)

```

The function `fitDLM` returns a list with two elements. The first element `results` is a data frame containing the original simulated data plus Kalman filtered and smoothed estimates of alpha and beta over time. The second is the AICc of the model though note that there are unanswered questions about whether this is calculated appropriately.  

And then we can plot the simulated stock-recruitment data (with points color-coded by year) and compare the true and estimated values of alpha (intercept) and beta (slope). 

```{r plotDLM, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4}
# first a little housekeeping
library(ggpubr)

spwn_range<-seq(0,max(dlm_model$results$spwn)*1.2,length.out=100)
r_pred <- matrix(NA,nrow=length(dlm_model$results$byr), ncol=length(spwn_range))


for(i in 1:length(dlm_model$results$byr)){
  r_pred[i,]<-exp(median(dlm_model$results$lnalpha[i]))*spwn_range*exp(dlm_model$results$beta[i]*spwn_range)
}

rownames(r_pred)<-dlm_model$results$byr
colnames(r_pred)<-spwn_range

r_pred<-cbind(dlm_model$results$byr,r_pred)
colnames(r_pred)[1]<-c("byr")
sr_pred<-pivot_longer(data =as.data.frame(r_pred), cols=!byr, names_to="spwn",values_to="rec" )
sr_pred$spwn<-as.numeric(sr_pred$spwn)

max_spawn <- max(dlm_model$result$spwn)
max_rec <- max(dlm_model$result$spwn)

# create each panel for plot starting with spawner-recruitment relationship

a <- ggplot(data=dlm_model$results, aes(x = spwn, y = rec, colour=factor(byr)))+
  geom_point(size = 3) +
  coord_cartesian(xlim=c(0, max_spawn*1.2), ylim=c(0,max_rec*1.2)) +
  xlab("Spawners") +
  ylab("Recruits") +
  geom_line(data = sr_pred, aes(x = spwn, y = rec, colour=factor(byr)), size = 0.5) +
  scale_colour_viridis_d()+
  theme_bw() +
  theme(strip.text.x = element_text(size=8),
        axis.title = element_text(size=10),
        axis.text = element_text(size=7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        plot.margin=unit(c(0.5,0.5,0.5,0.5), units="lines"))

# next plot true and estimate alpha

b <- ggplot(data=dlm_model$results, aes(x = byr, y = lnalpha )) +
  geom_line( color="black", size = 1)+
  geom_ribbon(aes(ymin = lnalpha-lnalpha_se*2, ymax = lnalpha+lnalpha_se*2), 
              fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
  geom_line(aes(x = byr, y = alpha_true), color="red", size = 1)+
  ylab("alpha") +
  xlab("") +
  scale_y_continuous(position = "right", limits=c(0,3)) +
  theme_bw() +
  theme(strip.text.x = element_text(size=8),
        axis.title = element_text(size=10),
        axis.text = element_text(size=7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        plot.margin=unit(c(0.5,2.25,0.05,0.5), units="lines"))+
  annotate("text", x = c(30,30),
           y = c(2.7, 2.5),
           label = c("True", "Estimated"),
           color= c("red", "black"), 
           size=3,
           hjust=0) +
  annotate("segment", x = c(28,28),
           xend=c(29.5,29.5),
           y = c(2.7, 2.5),
           yend = c(2.7, 2.5),
           lty = c(1,1),
           color=c("red", "black"), 
           size=1)

# next plot true and estimate beta

c<- ggplot(data=dlm_model$results, aes(x = byr, y = beta)) +
  geom_line( color="black", size = 1)+
  geom_ribbon(aes(ymin = beta-beta_se*2, ymax = beta+beta_se*2), 
              fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
  geom_line(aes(x = byr, y = -beta_true), color="red", size = 1)+
  xlab("Brood year") +
  ylab("beta") +
  scale_y_continuous(position = "right", limits=c(-0.0002,-0.000004)) +
  theme_bw() +
  theme(strip.text.x = element_text(size=8),
        axis.title = element_text(size=10),
        axis.text = element_text(size=7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        plot.margin=unit(c(0.05,0.5,0.5,0.5), units="lines"))

# combine plots
g <- ggarrange(a, ggarrange(b,c, nrow =2),
  ncol=2)

g
```





























