---
title: "Example fitting and visualizing for simulation scenrios"
author: "Brendan Connors, Luke Warkentin and Robyn Forrest"
date: "February 4, 2021"
output:
  html_document: default
  pdf_document: default
---

This document illustrates how to read in (or simulate) some simulated some stock-recruitment data from the `TESAsamSim` R package, fit a linear state-space model (aka Dynamic Linear model) with potentially time-varying parameters to it using a wrapper for the `dlm` R package (see [here](https://cran.r-project.org/web/packages/dlm/index.html)), and then visualize results.  

First, load packages, and read in (or run) simulations.

[RF: Figure out how to do all this with purrr or furrr later]

```{r, echo=FALSE, warnings=FALSE, message=FALSE}

## Check if necessary packages are available and install if necessary
listOfPackages <- c("here", "parallel", "doParallel", "foreach", 
                    "tidyverse", "tictoc", "TESAsamSim")
newPackages <- listOfPackages[!(listOfPackages %in% 
                                  installed.packages()[ , "Package"])]
if (length(newPackages)) {
  install.packages(newPackages)
}
lapply(listOfPackages, require, character.only = TRUE)

# load wrapper function to fit dynamic linear models
source(here("dlm-wrapper.R"))

simPar <- read.csv(here("data", "IFCohoPars",
                          "SimPar.csv"),
                        stringsAsFactors = F)
 
scenNames <- unique(simPar$scenario)

#if you need to get the sim files run this code (takes 15 mins)
#source("runTESAsamSim.R")


```

Put subset of scenarios in the same list then convert to a dataframe. Then create dataframes for output. Here we are just focusing on six alternative time trends in alpha and/or beta under two rho (autocorrelation in recruitment deviation) scenarios.

```{r, message=FALSE,  warning=FALSE}

simData <- list()
simData[[1]] <- readRDS(here("outputs", "simData", "runTESA",
                             "stationary.RData"))$srDatout
simData[[2]] <- readRDS(here("outputs", "simData", "runTESA",
                             "incLinearProd.RData"))$srDatout
simData[[3]] <- readRDS(here("outputs", "simData", "runTESA",
                             "incLinearCap.RData"))$srDatout
simData[[4]] <- readRDS(here("outputs", "simData", "runTESA",
                             "incLinearProdCap.RData"))$srDatout
simData[[5]] <- readRDS(here("outputs", "simData", "runTESA",
                             "decLinearProd.RData"))$srDatout
simData[[6]] <- readRDS(here("outputs", "simData", "runTESA",
                             "decLinearCap.RData"))$srDatout
simData[[7]] <- readRDS(here("outputs", "simData", "runTESA",
                             "decLinearProdCap.RData"))$srDatout
simData[[8]] <- readRDS(here("outputs", "simData", "runTESA",
                             "stationary_ar.RData"))$srDatout
simData[[9]] <- readRDS(here("outputs", "simData", "runTESA",
                             "incLinearProd_ar.RData"))$srDatout
simData[[10]] <- readRDS(here("outputs", "simData", "runTESA",
                             "incLinearCap_ar.RData"))$srDatout
simData[[11]] <- readRDS(here("outputs", "simData", "runTESA",
                             "incLinearProdCap_ar.RData"))$srDatout
simData[[12]] <- readRDS(here("outputs", "simData", "runTESA",
                             "decLinearProd_ar.RData"))$srDatout
simData[[13]] <- readRDS(here("outputs", "simData", "runTESA",
                             "decLinearCap_ar.RData"))$srDatout
simData[[14]] <- readRDS(here("outputs", "simData", "runTESA",
                             "decLinearProdCap_ar.RData"))$srDatout

# scenario IDs for subset we are working with
trunc_scenNames <- c("stationary","incLinearProd","incLinearCap","incLinearProdCap","decLinearProd","decLinearCap","decLinearProdCap","stationary_ar",
                     "incLinearProd_ar", "incLinearCap_ar","incLinearProdCap_ar","decLinearProd_ar","decLinearCap_ar","decLinearProdCap_ar")

# filter for CU and rename cols
for(i in 1:14){
  simData[[i]] <- simData[[i]] %>%
    dplyr::filter(CU == 1) %>%
    mutate() %>%
    select(-CU) %>%
    rename(byr=year, spwn=spawners, rec=recruits, alpha_true=alpha, beta_true=beta) %>%
    mutate(scen = trunc_scenNames[i], alpha=99., beta=99., alpha_se=99., beta_se=99.) %>% # cols for output
    select(scen, everything())  #reorder cols
}

# convert to df with all scenarios
dlm_Out <- simData[[1]]
for(i in 2:14) dlm_Out <- rbind(dlm_Out, simData[[i]])

# Now make copies to cover our alternative scenarios
dlm_Out <- dlm_Out %>%
  filter(byr>17)
estNames <- c("1_Stat", "2_Alpha_vary", "3_Beta_vary", "4_Alpha_Beta_vary")
dlm_Out$AIC <- 9999
dlm_Out$BIC <- 9999
dlm_Out_stat <- dlm_Out_alpha <- dlm_Out_beta <- dlm_Out_alphabeta <- dlm_Out

```

Now fit the simulated data for each scenario with each of 4 possible Dynamic Linear Models. 

```{r, message=FALSE,  warning=FALSE}

iter <- unique(simData[[i]]$iteration)
nsc <- length(trunc_scenNames)
for(j in 1:nsc){
  for(i in seq_along(iter)){

    dat <- dlm_Out %>%
      dplyr::filter(scen == trunc_scenNames[j], iteration==i) %>%
      select(-c(alpha,beta,alpha_se,beta_se)) #need to remove these for fitting

    # alpha and beta fixed in estimation model
    dlm_model_stat <- fitDLM(data = dat,
                             alpha_vary = FALSE,
                             beta_vary = FALSE)

    dlm_Out_stat[which(dlm_Out_stat$scen==trunc_scenNames[j] & dlm_Out_stat$iteration==i),10:13] <- dlm_model_stat$results[,12:15]
    dlm_Out_stat[which(dlm_Out_stat$scen==trunc_scenNames[j] & dlm_Out_stat$iteration==i),14] <- dlm_model_stat$AICc
    dlm_Out_stat[which(dlm_Out_stat$scen==trunc_scenNames[j] & dlm_Out_stat$iteration==i),15] <- dlm_model_stat$BIC

    # alpha varies in estimation model
    dlm_model_alpha <- fitDLM(data = dat,
                              alpha_vary = TRUE,
                              beta_vary = FALSE)

    dlm_Out_alpha[which(dlm_Out_alpha$scen==trunc_scenNames[j] & dlm_Out_alpha$iteration==i),10:13] <- dlm_model_alpha$results[,12:15]
    dlm_Out_alpha[which(dlm_Out_stat$scen==trunc_scenNames[j] & dlm_Out_alpha$iteration==i),14] <- dlm_model_alpha$AICc
    dlm_Out_alpha[which(dlm_Out_stat$scen==trunc_scenNames[j] & dlm_Out_alpha$iteration==i),15] <- dlm_model_alpha$BIC

    # beta varies in estimation model
    dlm_model_beta <- fitDLM(data = dat,
                             alpha_vary = FALSE,
                             beta_vary = TRUE)

    dlm_Out_beta[which(dlm_Out_beta$scen==trunc_scenNames[j] & dlm_Out_beta$iteration==i),10:13] <- dlm_model_beta$results[,12:15]
    dlm_Out_beta[which(dlm_Out_beta$scen==trunc_scenNames[j] & dlm_Out_beta$iteration==i),14] <- dlm_model_beta$AICc
    dlm_Out_beta[which(dlm_Out_beta$scen==trunc_scenNames[j] & dlm_Out_beta$iteration==i),15] <- dlm_model_beta$BIC

    # alpha and beta vary in estimation model
    dlm_model_alphabeta <- fitDLM(data = dat,
                                  alpha_vary = TRUE,
                                  beta_vary = TRUE)

    dlm_Out_alphabeta[which(dlm_Out_alphabeta$scen==trunc_scenNames[j] & dlm_Out_alphabeta$iteration==i),10:13] <- dlm_model_alphabeta$results[,12:15]
    dlm_Out_alphabeta[which(dlm_Out_alphabeta$scen==trunc_scenNames[j] & dlm_Out_alphabeta$iteration==i),14] <- dlm_model_alphabeta$AICc
    dlm_Out_alphabeta[which(dlm_Out_alphabeta$scen==trunc_scenNames[j] & dlm_Out_alphabeta$iteration==i),15] <- dlm_model_alphabeta$BIC

  }
}


# Now append the estimation model name to each dataframe
dlm_Out_stat <- dlm_Out_stat %>%
  mutate(estModel = estNames[1])
dlm_Out_alpha <- dlm_Out_alpha %>%
  mutate(estModel = estNames[2])
dlm_Out_beta <- dlm_Out_beta %>%
  mutate(estModel = estNames[3])
dlm_Out_alphabeta <- dlm_Out_alphabeta %>%
  mutate(estModel = estNames[4])

```

Now get all the outputs into one giant dataframe.

```{r, message=FALSE,  warning=FALSE, results="hide"}
dlm_out_all_combo <- rbind(dlm_Out_stat, dlm_Out_alpha, dlm_Out_beta, dlm_Out_alphabeta)

```

Next, lets calculate bias by parameter and magnitude of rho.

```{r, message=FALSE,  warning=FALSE}

pms <- dlm_out_all_combo %>%
  group_by(scen,estModel, iteration) %>%
  dplyr::summarize(
    alpha_mpb=mean((alpha_true-alpha)/alpha_true)*100,
    beta_mpb=mean((beta_true-(abs(beta)))/beta_true)*100) %>%
  pivot_longer(alpha_mpb:beta_mpb,names_to="parameter",values_to="mpb")

```

And then plot bias by estimation model, parameter and rho.
```{r, message=FALSE,  warning=FALSE}
# reorder scenarios
pms$scen <- factor(pms$scen, levels = c("stationary", "stationary_ar","incLinearProd","incLinearProd_ar","incLinearCap",
                                        "incLinearCap_ar", "incLinearProdCap","incLinearProdCap_ar","decLinearProd",
                                        "decLinearProd_ar","decLinearCap","decLinearCap_ar","decLinearProdCap", "decLinearProdCap_ar"))

# rename some variables for plotting
pms_rho <- pms %>%
  mutate(rho=case_when(scen=="stationary"~"rho = 0",
                       scen=="incLinearProd"~"rho = 0",
                       scen=="incLinearCap"~"rho = 0",
                       scen=="incLinearProdCap"~"rho = 0",
                       scen=="decLinearProd"~"rho = 0",
                       scen=="decLinearCap"~"rho = 0",
                       scen=="decLinearProdCap"~"rho = 0",
                       scen=="stationary_ar"~"rho = 0.4",
                       scen=="incLinearProd_ar"~"rho = 0.4",
                       scen=="incLinearCap_ar"~"rho = 0.4",
                       scen=="incLinearProdCap_ar"~"rho = 0.4",
                       scen=="decLinearProd_ar"~"rho = 0.4",
                       scen=="decLinearCap_ar"~"rho = 0.4",
                       scen=="decLinearProdCap_ar"~"rho = 0.4"))%>%
  mutate(scenMer=case_when(scen=="stationary"~"stationary",
                           scen=="incLinearProd"~"incLinearProd",
                           scen=="incLinearCap"~"incLinearCap",
                           scen=="incLinearProdCap"~"incLinearProdCap",
                           scen=="decLinearProd"~"decLinearProd",
                           scen=="decLinearCap"~"decLinearCap",
                           scen=="decLinearProdCap"~"decLinearProdCap",
                           scen=="stationary_ar"~"stationary",
                           scen=="incLinearProd_ar"~"incLinearProd",
                           scen=="incLinearCap_ar"~"incLinearCap",
                           scen=="incLinearProdCap_ar"~"incLinearProdCap",
                           scen=="decLinearProd_ar"~"decLinearProd",
                           scen=="decLinearCap_ar"~"decLinearCap",
                           scen=="decLinearProdCap_ar"~"decLinearProdCap")) %>%
  mutate(parameter2=case_when(parameter=="alpha_mpb"~"alpha",
                              parameter=="beta_mpb"~"beta"))

ggplot(data=pms_rho ,aes(x=scenMer, y=mpb, fill=estModel))+
  geom_boxplot(outlier.shape = NA)+
  xlab("Scenario") +
  ylab("Mean % bias") +
  coord_cartesian(ylim=c(-100,100))+
  scale_fill_viridis_d(labels=c("Stationary", "Alpha", "Beta", "Alpha & Beta"),name="Estimation\n model")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  geom_hline(yintercept=0, linetype="dashed", color = "red")+
  facet_grid(rho~parameter2)

ggsave(here("outputs", "simData", "runTESA", "bias-plot.jpeg"),height=6,width=8)

```

Next let's create a confusion matrix for a subset of scenarios with rho = 0. First we need to calculate type 1 and 2 errors.
 
```{r,  message=FALSE,  warning=FALSE}
conf_scenNames<-c("stationary","incLinearProd","incLinearCap","incLinearProdCap")
confusionOut <-dlm_out_all_combo %>%
  filter(scen==conf_scenNames)%>%
  group_by(scen,iteration,estModel)%>%
  filter(row_number()==1)

confusion<-matrix(0,400,6)
confusion<-as.data.frame(confusion)
confusion[,1]<-rep(unique(confusionOut$scen),each=100)
confusion[,2]<-rep(1:100,times=4)
colnames(confusion)<-c("scenario","MC","stationary","alpha","beta","alphabeta")
colnames(confusion)<-c("scenario","MC","stationary","alpha","beta","alphabeta")
confusionAIC <- confusion
confusionBIC <- confusion

  for(j in 1:4){
    for(i in unique(confusionOut$iteration)){
      AIC <- confusionOut[which(confusionOut$scen==conf_scenNames[j] & confusionOut$iteration==i),14]
      best_mod_AICc <- which.min(unlist(AIC))[1]
      BIC <- confusionOut[which(confusionOut$scen==conf_scenNames[j] & confusionOut$iteration==i),15]
      best_mod_BIC <- which.min(unlist(BIC))
      confusionAIC[which(confusionAIC$scen==conf_scenNames[j]& confusionAIC$MC==i),2+best_mod_AICc]<-1
      confusionBIC[which(confusionBIC$scen==conf_scenNames[j] & confusionBIC$MC==i),2+best_mod_BIC]<-1
    }
  }

confusionAIC$InfC <- rep("AIC",dim(confusionAIC)[1])
confusionBIC$InfC <- rep("BIC",dim(confusionBIC)[1])
confusion <- rbind(confusionAIC, confusionBIC)
confusion$InfC <- as.factor(confusion$InfC)


```

Then we can plot.

```{r, message=FALSE,  warning=FALSE}
confusionLong <- confusion %>%
  group_by(InfC,scenario) %>%
  dplyr::summarize(
    pStation=sum(stationary)/100,
    pAlpha=sum(alpha)/100,
    pBeta=sum(beta)/100,
    pAlphaBeta=sum(alphabeta)/100) %>%
  pivot_longer(pStation:pAlphaBeta,names_to="estimated",values_to="pMC") %>%
  mutate(trueNum=case_when(scenario=="stationary"~1,
                           scenario=="incLinearProd"~2,
                           scenario=="incLinearCap"~3,
                           scenario=="incLinearProdCap"~4)) %>%
  mutate(estNum=case_when(estimated=="pStation"~1,
                          estimated=="pAlpha"~2,
                          estimated=="pBeta"~3,
                          estimated=="pAlphaBeta"~4))

ggplot(data=confusionLong,aes(x=trueNum, y=estNum, fill=pMC))+
  geom_tile(aes(fill = pMC))+
  scale_fill_viridis_c(limits=c(0,1), name="Proportion\n of trials")+
  facet_wrap(~InfC)+
  xlab("True") +
  ylab("Estimated") +
  geom_text(aes(label = round(pMC, 2)), size = 3, color="white")+
  scale_x_continuous(labels=c("Stationary", "Alpha vary", "Beta vary", "Both vary"), breaks=c(1,2,3,4))+
  scale_y_continuous(labels=c("Stationary", "Alpha vary", "Beta vary", "Both vary"), breaks=c(1,2,3,4))+
  theme_bw()

ggsave(here("outputs", "simData", "runTESA", "confusion-matrix-rho=0.jpeg"),height=3.5,width=8)

```

How about another confusion matrix for a subset of scenarios with rho = 0.4? 
 
```{r}
conf_scenNames<-c("stationary_ar","incLinearProd_ar","incLinearCap_ar","incLinearProdCap_ar")

confusionOut <-dlm_out_all_combo %>%
  filter(scen==conf_scenNames)%>%
  group_by(scen,iteration,estModel)%>%
  filter(row_number()==1)

confusion<-matrix(0,400,6)
confusion<-as.data.frame(confusion)
confusion[,1]<-rep(unique(confusionOut$scen),each=100)
confusion[,2]<-rep(1:100,times=4)
colnames(confusion)<-c("scenario","MC","stationary","alpha","beta","alphabeta")
colnames(confusion)<-c("scenario","MC","stationary","alpha","beta","alphabeta")
confusionAIC <- confusion
confusionBIC <- confusion

  for(j in 1:4){
    for(i in unique(confusionOut$iteration)){
      AIC <- confusionOut[which(confusionOut$scen==conf_scenNames[j] & confusionOut$iteration==i),14]
      best_mod_AICc <- which.min(unlist(AIC))[1]
      BIC <- confusionOut[which(confusionOut$scen==conf_scenNames[j] & confusionOut$iteration==i),15]
      best_mod_BIC <- which.min(unlist(BIC))
      confusionAIC[which(confusionAIC$scen==conf_scenNames[j]& confusionAIC$MC==i),2+best_mod_AICc]<-1
      confusionBIC[which(confusionBIC$scen==conf_scenNames[j] & confusionBIC$MC==i),2+best_mod_BIC]<-1
    }
  }

confusionAIC$InfC <- rep("AIC",dim(confusionAIC)[1])
confusionBIC$InfC <- rep("BIC",dim(confusionBIC)[1])
confusion <- rbind(confusionAIC, confusionBIC)
confusion$InfC <- as.factor(confusion$InfC)


```

And then plot.

```{r}
confusionLong <- confusion %>%
  group_by(InfC,scenario) %>%
  dplyr::summarize(
    pStation=sum(stationary)/100,
    pAlpha=sum(alpha)/100,
    pBeta=sum(beta)/100,
    pAlphaBeta=sum(alphabeta)/100) %>%
  pivot_longer(pStation:pAlphaBeta,names_to="estimated",values_to="pMC") %>%
  mutate(trueNum=case_when(scenario=="stationary_ar"~1,
                           scenario=="incLinearProd_ar"~2,
                           scenario=="incLinearCap_ar"~3,
                           scenario=="incLinearProdCap_ar"~4)) %>%
  mutate(estNum=case_when(estimated=="pStation"~1,
                          estimated=="pAlpha"~2,
                          estimated=="pBeta"~3,
                          estimated=="pAlphaBeta"~4))

ggplot(data=confusionLong,aes(x=trueNum, y=estNum, fill=pMC))+
  geom_tile(aes(fill = pMC))+
  scale_fill_viridis_c(limits=c(0,1), name="Proportion\n of trials")+
  facet_wrap(~InfC)+
  xlab("True") +
  ylab("Estimated") +
  geom_text(aes(label = round(pMC, 2)), size = 3, color="white")+
  scale_x_continuous(labels=c("Stationary", "Alpha vary", "Beta vary", "Both vary"), breaks=c(1,2,3,4))+
  scale_y_continuous(labels=c("Stationary", "Alpha vary", "Beta vary", "Both vary"), breaks=c(1,2,3,4))+
  theme_bw()

ggsave(here("outputs", "simData", "runTESA", "confusion-matrix-rho=0.4.jpeg"),height=3.5,width=8)

```
